---
title: "Fast predictions from glmmTMB models"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This repository contains a single function, `predict_delta_glmmTMB`, which returns predicitons and confidence intervals for model predictions fitted with glmmTMB. Confidence intervals are computed using the Delta method. The function is much faster and uses less memory than `ggeffects::ggpredict`, but does not perform any checks.

## Example

1. Fake data

```{r cars}
set.seed(2)
n = 100

# covariates
data = replicate(2, rnorm(n, 0,1))

# parameters
b = c(0.5, 0.2)
sd = 0.2

# response
y = 0 + data %*% b + rnorm(n, sd = sd)
data = data.frame(y = y, data)

```


2. Fit model 

```{r, echo = FALSE, results=FALSE, message=FALSE}
library(glmmTMB)
m <- glmmTMB(y ~X1 + X2, data = data)
```


3. Generate predictions

```{r}
source('glmmTMB_predict_delta.R')

#newdata
newdata <- expand.grid(X1 = seq(-1.5, 1.5, length.out = 100),
                       X2 = c(-1.5, 0, 1.5))


pred <- predict_delta_glmmTMB(m, newdata)
head(pred)

```


4. Create plot

```{r, message=FALSE}
library(ggplot2)
(pred_plot <- ggplot(pred, aes(x = X1, y = fit)) +
  geom_line(aes(colour = as.factor(X2))) +
  geom_ribbon(aes(fill = as.factor(X2), ymin = conf_low, ymax = conf_high), alpha = 0.2) + 
  theme_minimal() + 
  theme(aspect.ratio = 1,
        legend.position = 'inside',
        legend.position.inside = c(0.8, 0.2)))
```


5. Compare to `ggeffects::ggpredict`

```{r}

library(ggeffects)
# generate predictions
pred_ggpredict <- data.frame(ggpredict(m, newdata))

(pred_plot_ggeffects <- ggplot(pred_ggpredict, aes(x = x, y = predicted)) +
  geom_line(aes(colour = group)) +
  geom_ribbon(aes(fill = group, ymin = conf.low, ymax = conf.high), alpha = 0.2) + 
  theme_minimal() + 
  theme(aspect.ratio = 1,
        legend.position = 'inside',
        legend.position.inside = c(0.8, 0.2))) 
```



